Reliable error estimation for Sobol' indices

### 2016-08-30 article/template.tex
changes L., corrections C.P

\newcommand{\fudge}{\fC}
\newcommand{\cc}{\mathcal{C}}
\newcommand{\dtf}{\textit{\doubletilde{f}}}
\newcommand{\cube}{[0,1)^d}
\newcommand{\rf}{\mathring{f}}
\newcommand{\rnu}{\mathring{\nu}}
\newcommand{\natm}{\naturals_{0,m}}
\newcommand{\wcS}{\widecheck{S}}
\newcommand{\tol}{\text{tol}}
\newcommand{\e}{\text{e}}
%\newcommand{\bvec}[1]{\boldsymbol{#1}}
\newcommand{\vect}[1]{\boldsymbol{\mathbf{#1}}}
\newcommand{\vx}{\vect{x}}
\newcommand{\vI}{\vect{I}}
\newcommand{\vk}{\vect{k}}
\newcommand{\vw}{\vect{w}}
\newcommand{\vz}{\vect{z}}
\newcommand{\vt}{\vect{t}}
\newcommand{\dif}{\mathsf{d}}
\newcommand{\hf}{\hat{f}}
\newcommand{\hS}{\widehat{S}}
\newcommand{\tS}{\widetilde{S}}
\newcommand{\tf}{\tilde{f}}
\newcommand{\fC}{\mathfrak{C}}
\newcommand{\homega}{\widehat{\omega}}
\newcommand{\wcomega}{\mathring{\omega}}
\newcommand{\vzero}{\vect{0}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\naturals}{\mathbb{N}}
\newcommand{\ip}[3][{}]{\ensuremath{\left \langle #2, #3 \right \rangle_{#1}}}
\newcommand\iid{\stackrel{{\rm iid}}{\sim}}

@@

Firstly, section \ref{sec:2} introduces Sobol' indices and reviews both Saltelli's strategy to estimate first-order and total effect Sobol' indices, and the replication procedure. Our main contribution is detailed in  Section \ref{sec:3}. There we review the construction of the error bound for the estimation of integrals based on digital nets, and then generalize it for Sobol' indices. Section \ref{sec:4} is devoted to analyze the cost in terms of model evaluations of our sequential estimation algorithm. It combines the error bound in Section \ref{sec:3} and one of the two estimation procedures of Section \ref{sec:2}. We also discuss a potential improvement to estimate small first-order indices according to \cite{Owen}. Finally, examples and illustrations of our procedure are presented in Section \ref{sec:5}.

@@

\subsection{Definition of Sobol' indices}
\label{sec:2.1}
%We adopt the same notations introduced by Owen in \cite{Owen}. 
Denote by $\vx=(x_1,\dots,x_d)$ the vector of inputs of $f$ and $\mathcal{D}=\{1,\dots,d\}$ the set of dimensions indexes. %The uncertainty on $\vx$ is modeled by a random vector that we suppose uniformly distributed on $\cube$.
Let $u$ be a subset of $\mathcal{D}$, $-u$ its complement and $|u|$ its cardinality. $\vx_u$ represents a point in $[0,1)^{|u|}$ with components $x_j, j \in u$. Given two points $\vx$ and $\vx'$, the hybrid point $\vw=(\vx_u:{\vx'}_{-u})$ is defined as $w_j=x_j$ if $j \in u$ and $w_j={x'}_j$ if $j \notin u$.

We assume that $f \in \mathcal{L}^2\left([0,1]^d\right)$ and denote by $\mu$ and $\sigma^2$ the mean $\int_{\cube}f(\vx)d{\vx}$ and variance $\int_{\cube}(f(\vx)-\mu)^2d{\vx}$ of $f$.
The uncertainty on $\vx$ is modeled as a uniform random variable, namely $\vx \iid \mathcal{U}\left([0,1)^d\right)$. The Hoeffding decomposition \cite{Hoeffding,Sobol'} of $f$ is:
\begin{equation}
f(\vx)=\mu+\sum \limits_{u \subseteq \mathcal{D}, u \neq \varnothing} f_u(\vx),
\label{anova}
\end{equation}
where:
\[f_u(\vx)= \int_{[0,1)^{|u|}} f(\vx) d{\vx}_{-u} - \sum \limits_{v \subset u} f_v(\vx).\]
Due to orthogonality, applying the variance to equation \eqref{anova} leads to the variance decomposition of $f$:
\[ \sigma^2 = \sum \limits_{u \subseteq \mathcal{D}, u \neq \varnothing} \sigma_u^2, \ \text{ with } \ \sigma_u^2=\int_{[0,1)^{|u|}} f_u(\vx)^2 d{\vx_u}.\]
From this decomposition, one can express the following two quantities:
\[\underline{\tau}_u^2 = \sum \limits_{v \subseteq u} \sigma_v^2, \qquad
\ov{$\tau$}_u^2 = \sum \limits_{v \cap u \neq \varnothing} \sigma_v^2, \qquad u \subsetneq \mathcal{D}.\]

@@

The problem of interest is the evaluation of first-order and total effect Sobol' indices of order 1. The computation of the normalized indices is performed based on the following integral formulas for their numerators:
\begin{align}
\label{first.order}
\underline{\tau}_u^2  &= \int_{[0,1)^{2d-1}} \left(f(\vx)-
f(\vx_u:{\vx'}_{-u})\right)f(\vx')d\vx d{\vx'_{-u}}, \\
\label{total.effect}
\ov{$\tau$}_u^2 &= \frac{1}{2}\int_{[0,1)^{d+1}}(f(\vx')-f(\vx_u:{\vx'}_{-u}))^2 d\vx_u d{\vx'}, \ u \in \mathcal{D},
\end{align}
while the variance of $f$ is evaluated as:
\begin{equation}
\sigma^2 = \int_{[0,1)^{d}} f(\vx)^2d{\vx} - \mu^2, \text{ with } \ \mu = \int_{[0,1)^{d}} f(\vx) d{\vx}.
\label{eq.sigma.mu}
\end{equation} 
Usually the complexity of $f$ causes the computation of integrals \eqref{first.order}, (\ref{total.effect}) and (\ref{eq.sigma.mu}) to be intractable. In such cases, one can estimate these quantities instead.

\subsection{Estimation of Sobol' indices}
\label{sec:2.2}
In this section we review two Monte Carlo procedures for the estimation of Sobol' indices. A design is a point set $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ where each point is obtained by sampling each variable $x_j$ $n$ times. Each row of the design is a point $\vx_i$ in $[0,1)^d$ and each column of the design refers to a variable $x_j$. Consider $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P'}=\{{\vx'}_i\}_{i=0}^{n-1}$ two designs where $(\vx_i,{\vx'}_i) \iid [0,1)^{2d}$. One way to estimate the two quantities (\ref{first.order}) and (\ref{total.effect}) is via:
\begin{align}
\label{first.order.est}
\widehat{\underline{\tau}_u^2} &= \frac{1}{n} \sum \limits_{i=0}^{n-1} \left(f(\vx_i)-f(\vx_{i,u}:{\vx'}_{i,-u})\right)f(\vx'_i),\\
\label{total.effect.est}
\widehat{\ov{$\tau$}_u^2} &= \frac{1}{2n} \sum \limits_{i=0}^{n-1} (f({\vx'}_i) - f(\vx_{i,u}:{\vx'}_{i,-u}))^2, \qquad u \in \mathcal{D},
\end{align}
using for $\sigma^2$:
\begin{equation}
 \widehat{\sigma}^2 = \frac{1}{n} \sum \limits_{i=0}^{n-1} f(\vx_i)^2 - \widehat{\mu}^2, \text{ with } \ \widehat{\mu} =  \frac{1}{n} \sum \limits_{i=0}^{n-1} f(\vx_i).
\label{mu.est}
\end{equation}
Then, the normalized Sobol' indices estimators are:
\begin{equation}
\widehat{\underline{S}_u} = \widehat{\underline{\tau}_u^2} / \widehat{\sigma}^2, \qquad
\widehat{\overline{S}_u} = \widehat{\overline{\tau}_u^2} / \widehat{\sigma}^2.
\label{common.sobol.est}
\end{equation}
The estimation of a single pair ($\widehat{\underline{S}_u}$, $\widehat{\overline{S}_u}$) requires $3n$ evaluations of the model $f$. Thus, for first order indices, one would need $3nd$ model evaluations. Using a combinatorial formalism, in \cite{Saltelli}, Saltelli proposes the following estimation strategy:
\begin{theorem}
\label{saltelli.theorem}
The $d+2$ designs $\{\vx_{i,u},{\vx'}_{i,-u}\}_{i=0}^{n-1}$ constructed for $u \in \{\varnothing,\{1\},\dots,$ $\{d\},\mathcal{D}\}$ allows to estimate all first-order and total effect Sobol' indices of order 1 at a cost of $n(d+2)$ evaluations of the model.
\end{theorem}
The $d+2$ designs of Theorem \ref{saltelli.theorem} are obtained by substituting columns of $\mathcal{P}$ for columns of $\mathcal{P}'$ accordingly to $u$. Indeed, there is no need of reevaluating $f({\vx}_i$ and $f({\vx'}_i$ for each $u$. While elegant, this approach requires a number of model evaluations that grows linearly with respect to the input space dimension.

@@

More precisely, let $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P}'=\{{\vx'}_i\}_{i=0}^{n-1}$ be two replicated designs of order $1$. Denote by $\{f(\vx_i)\}_{i=0}^{n-1}$ and $\{f({\vx'}_i)\}_{i=0}^{n-1}$ the two sets of model evaluations obtained with $\mathcal{P}$ and $\mathcal{P}'$. From Definition \ref{rep.designs}, ${\vx'}_{\pi_u(i),u}={\vx}_{i,u}$. Then, for all $i \in \{0,\dots,n-1\}$:
\begin{align*}
\pi_u(f({\vx'}_i))&= f({\vx'}_{\pi_u(i)}), \\
&= f(\vx'_{\pi_u(i),u}:{\vx'}_{\pi_u(i),-u}), \\
&= f(\vx_{i,u}:{\vx'}_{\pi_u(i),-u}).
\end{align*}
Hence, each $\underline{\tau}^2_u$ can be estimated via formula (\ref{first.order.est}) with \\ $\pi_u(f({\vx'}_i))$ in place of $f(\vx_{i,u}:{\vx'}_{i,-u})$ without requiring further model evaluations. This estimation method has been deeply studied and generalized in Tissot et al. \cite{Tissot} to the case of closed second-order indices. In the following we refer to this method as replication procedure.

\subsection{Towards a reliable estimation}
\label{sec:2.3}
The aim of this paper is to propose a sequential procedure to estimate first-order and total effect Sobol' indices. Generally, in most sequential procedures that estimate Sobol' indices, the stopping criterion is a quantity of interest built directly upon the estimates. Such stopping criteria often involve hyper-parameters hard to tweak but above all fail to guarantee any error bound on the estimates.

@@

In order to relate the group structure of $\mathcal{P}_m$ with the integration error, we introduce the \emph{dual net} which establishes the relationship between any digital net and the \emph{wavenumber} space $\mathbb{N}_0^d$. Hence, a dual net is
\[
\mathcal{P}_m^\perp=\{\vk\in\mathbb{N}_0^d:\ip{\vx}{\vk}=0,\, \vx\in\mathcal{P}_m\},
\]
and inherits the same embedded structure as for the digital nets,
\begin{equation}\label{dual_net_structure}
\mathcal{P}_0^\perp=\mathbb{N}_0^d\supset\dots\supset\mathcal{P}_\infty=\{\vect{0}\}.
\end{equation}
As shown in \cite[Sec. 3]{HicJim}, the group structure of the digital nets guarantees the property below affecting any Walsh basis $\phi_{\vk}(\vx)$,
\begin{equation}\label{basis_integ_prop}
\frac{1}{b^m}\sum_{\vx\in\mathcal{P}_m}\phi_{\vk}(\vx)=
\begin{cases}
1,\quad \vk\in\mathcal{P}_m^\perp, \\
0,\quad \vk\notin\mathcal{P}_m^\perp.
\end{cases}
\end{equation}
Therefore, considering the Walsh decomposition for any $f\in L^2([0,1]^d)$: \[
f(\vx)=\sum_{\vk\in\naturals_0^d}\hf_{\vk}\varphi_{\vk}(\vx), \]
and,
\[ I(f) = \int_{\cube} f(\vx)\,d\vx,\]
property \eqref{basis_integ_prop} leads to
\begin{align*}
\absolute{I(f) - \frac{1}{b^m}\sum_{\vx\in\mathcal{P}_m}f(\vx)} &=
\absolute{\hf_{\vzero} - \frac{1}{b^m}\sum_{\vx\in\mathcal{P}_m}\sum_{\vk\in\naturals_0^d}\hf_{\vk}\varphi_{\vk}(\vx)}, \\
&= \absolute{\sum_{\vk\in\mathcal{P}_m^\perp\setminus\{\vzero\}}\hf_{\vk}} \\
& \leq \sum_{\vk\in\mathcal{P}_m^\perp\setminus\{\vzero\}}\absolute{\hf_{\vk}}.
\end{align*}
Based on the size of $\absolute{\hf_{\vk}}$ and the structure of the dual nets \eqref{dual_net_structure}, in \cite[Sec. 4.1]{HicJim} we proposed an ordering of the wave-numbers $\vk(\cdot):\naturals_0\rightarrow\naturals_0^d$. To simplify notation, $\hf_{\kappa}:=\hf_{\vk(\kappa)}$. This mapping leads to the following error bound,
\begin{equation}
\absolute{I(f) - \frac{1}{b^m}\sum_{\vx\in\mathcal{P}_m}f(\vx)} \le \sum_{\lambda=1}^{\infty} \left \lvert \hf_{\lambda b^m}\right \rvert.
\end{equation}
However, because the knowledge of the Walsh coefficients $\hf_{\kappa}$ is not assumed, we will estimate them using the precomputed function values instead, and refer to them as $\tf_{m,\kappa}$.

For $\ell,m \in \mathbb{N}_0$ and $\ell \le m$ we introduce the following notation,
\begin{align*}
S_m(f) &=  \sum_{\kappa=\left \lfloor b^{m-1} \right \rfloor}^{b^{m}-1} \absolute{\hf_{\kappa}}, \ 
\hS_{\ell,m}(f)  = \sum_{\kappa=\left \lfloor b^{\ell-1} \right \rfloor}^{b^{\ell}-1} \sum_{\lambda=1}^{\infty} \absolute{ \hf_{\kappa+\lambda b^{m}}}, \\
\wcS_m(f)&=\hS_{0,m}(f) + \cdots + \hS_{m,m}(f)=
\sum_{\kappa=b^{m}}^{\infty} \absolute{\hf_{\kappa}},\\ 
\tS_{\ell,m}(f) &= \sum_{\kappa=\left \lfloor b^{\ell-1}\right \rfloor}^{b^{\ell}-1} \absolute{\tf_{m,\kappa}}.
\end{align*}
Finally, we define the set of functions $\cc$,
\begin{multline} \label{conecond}
\cc:=\{f \in L^2(\cube) : \hS_{\ell,m}(f) \le \homega(m-\ell) \wcS_m(f),\ \ \ell \le m, \\
\wcS_m(f) \le \wcomega(m-\ell) S_{\ell}(f),\ \  \ell_* \le \ell \le m\}.
\end{multline}
for $\ell_* \in \mathbb{N}$, $\homega$ and $\wcomega$ two non-negative valued functions with $\lim_{m \to \infty} \wcomega(m) = 0$ and such that $\homega(r)\wcomega(r)<1$ for some $r\in\mathbb{N}$.

Hence, in \cite[Sec. 4.2]{HicJim} we prove that for any $f\in\cc$,
\begin{equation}
\absolute{I(f) - \frac{1}{b^m}\sum_{\vx\in\mathcal{P}_m}f(\vx)}
\le \overbrace{\frac{\tS_{\ell,m}(f)\homega(m) \wcomega(m-\ell)}{1 - \homega(m-\ell) \wcomega(m-\ell)}}^{\epsilon_I:=}, \label{errbd}
\end{equation}
where one may increase $m$ until the error bound $\epsilon_I$ is small enough.

@@

We start be detailing our procedure in the form of an algorithm. Then, we discuss a possible improvement by considering a new estimator recently introduced in \cite{Owen} for the estimation of first-order indices.

@@

In both case we test if the respective error bounds $\varepsilon_{\underline{S}_u}$ and $\varepsilon_{\overline{S}_u}$ are lower than the tolerance $\epsilon$. For Variant $A$ the stopping criterion writes: $ \forall u \in \mathcal{D}, \varepsilon_{\underline{S}_u} \leq \epsilon \text{ and } \varepsilon_{\overline{S}_u} \leq \epsilon$,
for Variant $B$ it writes:
$ \forall u \in \mathcal{D}, \varepsilon_{\underline{S}_u} \leq \epsilon.$

@@

In pratice, one does not know which are the small Sobol' indices. To overcome this issue, we propose the following alternative for Variant $A$. If at the end of the first iteration, the estimate $\widehat{\underline{S}_u}$, $u \in \mathcal{D}$, is smaller than a specified threshold, then at the next iteration estimator (\ref{first.order.est}) is switched for ``Correlation 2" and a third Sobol' sequence $\mathcal{P''}_{m}=\{\vz_i\}_{i=0}^{2^m-1}$ is constructed to obtain the corresponding evaluations $f({\vz}_{i,u}:{\vx}_{i,-u})$.

>>>>

\newcommand{\fudge}{\fC}
\newcommand{\cc}{\mathcal{C}}
\newcommand{\dtf}{\textit{\doubletilde{f}}}
\newcommand{\cube}{[0,1]^d}
\newcommand{\rf}{\mathring{f}}
\newcommand{\rnu}{\mathring{\nu}}
\newcommand{\natm}{\naturals_{0,m}}
\newcommand{\wcS}{\widecheck{S}}
\newcommand{\tol}{\text{tol}}
\newcommand{\e}{\text{e}}
%\newcommand{\bvec}[1]{\boldsymbol{#1}}
\newcommand{\vect}[1]{\boldsymbol{\mathbf{#1}}}
\newcommand{\vx}{\vect{x}}
\newcommand{\vI}{\vect{I}}
\newcommand{\vk}{\vect{k}}
\newcommand{\vw}{\vect{w}}
\newcommand{\vz}{\vect{z}}
\newcommand{\vt}{\vect{t}}
\newcommand{\dif}{\mathsf{d}}
\newcommand{\hf}{\hat{f}}
\newcommand{\hS}{\widehat{S}}
\newcommand{\tS}{\widetilde{S}}
\newcommand{\tf}{\tilde{f}}
\newcommand{\fC}{\mathfrak{C}}
\newcommand{\homega}{\widehat{\omega}}
\newcommand{\wcomega}{\mathring{\omega}}
\newcommand{\vzero}{\vect{0}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\naturals}{\mathbb{N}}
\newcommand{\ip}[3][{}]{\ensuremath{\left \langle #2, #3 \right \rangle_{#1}}}
\newcommand\iid{\stackrel{{\rm iid}}{\sim}}

@@

Firstly, section \ref{sec:2} introduces Sobol' indices and reviews both Saltelli's strategy to estimate first-order and total effect Sobol' indices, and the replication procedure. Our main contribution is detailed in  Section \ref{sec:3}. There we review the construction of the error bound for the estimation of integrals based on digital nets, and then generalize it for Sobol' indices. Section \ref{sec:4} is devoted to analyze the cost in terms of model evaluations of our sequential estimation algorithm. It combines the error bound in Section \ref{sec:3} and either one of the two estimation procedures of Section \ref{sec:2}. We also discuss a potential improvement to estimate small first-order indices according to \cite{Owen}. Finally, examples and illustrations of our procedure are presented in Section \ref{sec:5}.

@@

\subsection{Definition of Sobol' indices}
\label{sec:2.1}
%We adopt the same notations introduced by Owen in \cite{Owen}. 
Denote by $\vx=(x_1,\dots,x_d)$ the vector of inputs of $f$ and $\mathcal{D}=\{1,\dots,d\}$ the set of dimensions indexes. %The uncertainty on $\vx$ is modeled by a random vector that we suppose uniformly distributed on $\cube$.
Let $u$ be a subset of $\mathcal{D}$, $-u$ its complement and $|u|$ its cardinality. $\vx_u$ represents a point in $[0,1)^{|u|}$ with components $x_j, j \in u$. Given two points $\vx$ and $\vx'$, the hybrid point $\vw=(\vx_u:{\vx'}_{-u})$ is defined as $w_j=x_j$ if $j \in u$ and $w_j={x'}_j$ if $j \notin u$. We assume that $f \in \mathbb{L}^2(\cube)$.% and denote by $\mu$ and $\sigma^2$ the mean $\int_{\cube}f(\vx)d{\vx}$ and variance $\int_{\cube}(f(\vx)-\mu)^2d{\vx}$ of $f$.

The uncertainty on $\vx$ is modeled by a uniform random vector, namely $\vx \iid \mathcal{U}(\cube)$. The Hoeffding decomposition \cite{Hoeffding,Sobol'} of $f$ is:
\begin{equation}
f(\vx)=f_{\varnothing}+\sum \limits_{u \subseteq \mathcal{D}, u \neq \varnothing} f_u(\vx),
\label{anova}
\end{equation}
where:
\begin{align*}
f_{\varnothing} &= \mathrm{E}[\vx]= \mu, \\
f_u(\vx) &= \int_{[0,1)^{|u|}} f(\vx) d{\vx}_{-u} - \sum \limits_{v \subset u} f_v(\vx).
\end{align*}
Due to orthogonality, applying the variance to equation \eqref{anova} leads to the variance decomposition of $f$:
\[ \sigma^2 = \mathrm{Var}[\vx] = \sum \limits_{u \subseteq \mathcal{D}, u \neq \varnothing} \sigma_u^2, \ \text{ with } \ \sigma_u^2=\int_{[0,1)^{|u|}} f_u(\vx)^2 d{\vx_u}.\]
From this decomposition, one can express the following two quantities:
\[\underline{\tau}_u^2 = \sum \limits_{v \subseteq u} \sigma_v^2, \qquad
\ov{$\tau$}_u^2 = \sum \limits_{v \cap u \neq \varnothing} \sigma_v^2, \qquad u \subsetneq \mathcal{D}.\]

@@

The problem of interest is the evaluation of first-order and total effect Sobol' indices. In our framework, we are only interested in single input indices. To this effect we restrain $u \in \mathcal{D}$. The computation of the normalized indices is performed based on the following integral formulas for their numerators:
\begin{align}
\label{first.order}
\underline{\tau}_u^2  &= \int_{[0,1)^{2d-1}} \left(f(\vx)-
f(\vx_u:{\vx'}_{-u})\right)f(\vx')d\vx d{\vx'_{-u}}, \\
\label{total.effect}
\ov{$\tau$}_u^2 &= \frac{1}{2}\int_{[0,1)^{d+1}}(f(\vx')-f(\vx_u:{\vx'}_{-u}))^2 d\vx_u d{\vx'}, \ u \in \mathcal{D},
\end{align}
while variance and mean of $f$ are evaluated by:
\begin{gather}
\begin{aligned}
\sigma^2 &= \int_{[0,1)^{d}} f(\vx)^2d{\vx} - \mu^2, \\
 \mu &= \int_{[0,1)^{d}} f(\vx) d{\vx},
\label{eq.sigma.mu}
\end{aligned}
\end{gather}
Usually the complexity of $f$ causes the computation of integrals \eqref{first.order}, (\ref{total.effect}) and (\ref{eq.sigma.mu}) to be intractable. In such cases, one can estimate these quantities instead.

\subsection{Estimation of Sobol' indices}
\label{sec:2.2}
In this section we review two Monte Carlo procedures for the estimation of Sobol' indices. A design is a point set $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ where each point is obtained by sampling each variable $x_j$ $n$ times. Each row of the design is a point $\vx_i$ in $[0,1)^d$ and each column of the design refers to a variable $x_j$. Consider $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P'}=\{{\vx'}_i\}_{i=0}^{n-1}$ two designs where $(\vx_i,{\vx'}_i) \iid [0,1)^{2d}$. One way to estimate the two quantities (\ref{first.order}) and (\ref{total.effect}) is via:
\begin{align}
\label{first.order.est}
\widehat{\underline{\tau}_u^2} &= \frac{1}{n} \sum \limits_{i=0}^{n-1} \left(f(\vx_i)-f(\vx_{i,u}:{\vx'}_{i,-u})\right)f(\vx'_i),\\
\label{total.effect.est}
\widehat{\ov{$\tau$}_u^2} &= \frac{1}{2n} \sum \limits_{i=0}^{n-1} (f({\vx'}_i) - f(\vx_{i,u}:{\vx'}_{i,-u}))^2, \qquad u \in \mathcal{D},
\end{align}
using for $\sigma^2$:
\begin{equation}
 \widehat{\sigma}^2 = \frac{1}{n} \sum \limits_{i=0}^{n-1} f(\vx_i)^2 - \widehat{\mu}^2, \text{ with } \ \widehat{\mu} =  \frac{1}{n} \sum \limits_{i=0}^{n-1} f(\vx_i).
\label{mu.est}
\end{equation}
Then, the Sobol' indices estimators are:
\begin{equation}
\widehat{\underline{S}}_u = \widehat{\underline{\tau}_u^2} / \widehat{\sigma}^2, \qquad
\widehat{\overline{S}}_u = \widehat{\overline{\tau}_u^2} / \widehat{\sigma}^2.
\label{common.sobol.est}
\end{equation}
The estimation of a single pair ($\widehat{\underline{S}}_u$, $\widehat{\overline{S}}_u$) requires $3n$ evaluations of the model $f$. Thus, for first order indices, one would need $3nd$ model evaluations. Using a combinatorial formalism, in \cite{Saltelli}, Saltelli proposes the following estimation strategy:
\begin{theorem}
\label{saltelli.theorem}
The $d+2$ designs $\{\vx_{i,u}:{\vx'}_{i,-u}\}_{i=0}^{n-1}$ constructed for $u \in \{\varnothing,\{1\},\dots,$ $\{d\},\mathcal{D}\}$ allows to estimate all first-order and total effect Sobol' indices of order 1 at a cost of $n(d+2)$ evaluations of the model.
\end{theorem}
The $d+2$ designs of Theorem \ref{saltelli.theorem} are obtained by substituting columns of $\mathcal{P}$ for columns of $\mathcal{P}'$ accordingly to $u$. Indeed, there is no need of reevaluating $f({\vx}_i)$ and $f({\vx'}_i)$ for each $u$. While elegant, this approach requires a number of model evaluations that grows linearly with respect to the input space dimension.

@@

More precisely, let $\mathcal{P}=\{\vx_i\}_{i=0}^{n-1}$ and $\mathcal{P}'=\{{\vx'}_i\}_{i=0}^{n-1}$ be two replicated designs of order $1$. Denote by $\{f(\vx_i)\}_{i=0}^{n-1}$ and $\{f({\vx'}_i)\}_{i=0}^{n-1}$ the two sets of model evaluations obtained with $\mathcal{P}$ and $\mathcal{P}'$. From Definition \ref{rep.designs}, ${\vx'}_{\pi_u(i),u}={\vx}_{i,u}$. We define:
\[\pi_u(f({\vx'}_i)) = f({\vx'}_{\pi_u(i)}), \qquad  i \in \{0,\dots,n-1\},\]
then remark:
\begin{align*}
\pi_u(f({\vx'}_i))&= f(\vx'_{\pi_u(i),u}:{\vx'}_{\pi_u(i),-u}), \\
&= f(\vx_{i,u}:{\vx'}_{\pi_u(i),-u}).
\end{align*}
Hence, each $\underline{\tau}^2_u$ can be estimated via formula (\ref{first.order.est}) with \\ $\pi_u(f({\vx'}_i))$ in place of $f(\vx_{i,u}:{\vx'}_{i,-u})$ without requiring further model evaluations. This estimation method has been deeply studied and generalized in Tissot et al. \cite{Tissot} to the case of closed second-order indices. In the following we refer to this method as replication procedure.

\subsection{Towards a reliable estimation}
\label{sec:2.3}
The aim of this paper is to propose a sequential procedure to estimate first-order and total effect Sobol' indices.  A practical problem concerning the estimation of these indices is how large to choose the number of evaluations to ensure that Sobol' estimates are accurate enough. Asymptotical results show that Sobol' estimates are normally distributed (\cite[Proposition 2.2]{Janon}, \cite[Proposition 3.5]{Tissot}). As a consequence, confidence intervals can be use.

An alternative to address this problematic is the use of sequential procedures. As examples, sequential versions of the replication procedure and McKay's procedure are respectively proposed in \cite{Gilquin.rec} and \cite{Tong}. However, in those two latter procedures, the stopping criterion is a quantity of interest purely empirical, built directly upon the estimates. Such stopping criteria often involve hyper-parameters difficult to tweak but more importantly fail to guarantee any error bound on the estimates.

@@

In order to relate the group structure of $\mathcal{P}_m$ with the integration error, we introduce the \emph{dual net} which establishes the relationship between any digital net and the \emph{wavenumber} space $\mathbb{N}_0^d$ of non-negative integers. Hence, a dual net is
\[
\mathcal{P}_m^\perp=\{\vk\in\mathbb{N}_0^d:\ip{\vx}{\vk}=0,\, \vx\in\mathcal{P}_m\},
\]
and inherits the same embedded structure as for the digital nets,
\begin{equation}\label{dual_net_structure}
\mathcal{P}_0^\perp=\mathbb{N}_0^d\supset\dots\supset\mathcal{P}_\infty^\perp=\{\vect{0}\}.
\end{equation}
As shown in \cite[Sec. 3]{HicJim}, the group structure of the digital nets guarantees the property below affecting any Walsh basis $\varphi_{\vk}(\vx)$,
\begin{equation}\label{basis_integ_prop}
\frac{1}{b^m}\sum_{\vx\in\mathcal{P}_m}\varphi_{\vk}(\vx)=
\begin{cases}
1,\quad \vk\in\mathcal{P}_m^\perp, \\
0,\quad \vk\notin\mathcal{P}_m^\perp.
\end{cases}
\end{equation}
Therefore, considering the Walsh decomposition for any $f\in \mathbb{L}^2([0,1]^d)$: \[
f(\vx)=\sum_{\vk\in\naturals_0^d}\hf_{\vk}\varphi_{\vk}(\vx), \]
and,
\[ I(f) = \int_{\cube} f(\vx)\,d\vx,\]
property \eqref{basis_integ_prop} leads to
\begin{align*}
\absolute{I(f) - \frac{1}{b^m}\sum_{\vx\in\mathcal{P}_m}f(\vx)} &=
\absolute{\hf_{\vzero} - \frac{1}{b^m}\sum_{\vx\in\mathcal{P}_m}\sum_{\vk\in\naturals_0^d}\hf_{\vk}\varphi_{\vk}(\vx)}, \\
&= \absolute{\sum_{\vk\in\mathcal{P}_m^\perp\setminus\{\vzero\}}\hf_{\vk}} \\
& \leq \sum_{\vk\in\mathcal{P}_m^\perp\setminus\{\vzero\}}\absolute{\hf_{\vk}}.
\end{align*}
Based on the size of $\absolute{\hf_{\vk}}$ and the structure of the dual nets \eqref{dual_net_structure}, in \cite[Sec. 4.1]{HicJim} we proposed an ordering of the wave-numbers $\vk(\cdot):\naturals_0\rightarrow\naturals_0^d$. To simplify notation, $\hf_{\kappa}:=\hf_{\vk(\kappa)}$. This mapping leads to the following error bound,
\begin{equation}
\absolute{I(f) - \frac{1}{b^m}\sum_{\vx\in\mathcal{P}_m}f(\vx)} \le \sum_{\lambda=1}^{\infty} \left \lvert \hf_{\lambda b^m}\right \rvert.
\end{equation}
However, because the knowledge of the Walsh coefficients $\hf_{\kappa}$ is not assumed, we will estimate them using the precomputed function values instead, and refer to them as $\tf_{m,\kappa}$.

For $\ell,m \in \mathbb{N}_0$ and $\ell \le m$ we introduce the following notation,
\begin{align*}
S_m(f) &=  \sum_{\kappa=\left \lfloor b^{m-1} \right \rfloor}^{b^{m}-1} \absolute{\hf_{\kappa}}, \ 
\hS_{\ell,m}(f)  = \sum_{\kappa=\left \lfloor b^{\ell-1} \right \rfloor}^{b^{\ell}-1} \sum_{\lambda=1}^{\infty} \absolute{ \hf_{\kappa+\lambda b^{m}}}, \\
\wcS_m(f)&=\hS_{0,m}(f) + \cdots + \hS_{m,m}(f)=
\sum_{\kappa=b^{m}}^{\infty} \absolute{\hf_{\kappa}},\\ 
\tS_{\ell,m}(f) &= \sum_{\kappa=\left \lfloor b^{\ell-1}\right \rfloor}^{b^{\ell}-1} \absolute{\tf_{m,\kappa}}.
\end{align*}
Finally, we define the set of functions $\cc$,
\begin{multline} \label{conecond}
\cc:=\{f \in \mathbb{L}^2(\cube) : \hS_{\ell,m}(f) \le \homega(m-\ell) \wcS_m(f),\ \ \ell \le m, \\
\wcS_m(f) \le \wcomega(m-\ell) S_{\ell}(f),\ \  \ell_* \le \ell \le m\}.
\end{multline}
for $\ell_* \in \mathbb{N}$, $\homega$ and $\wcomega$ two non-negative valued functions with $\lim_{m \to \infty} \wcomega(m) = 0$ and such that $\homega(r)\wcomega(r)<1$ for some $r\in\mathbb{N}$.

Hence, in \cite[Sec. 4.2]{HicJim} it is proved that for any $f\in\cc$,
\begin{equation}
\absolute{I(f) - \frac{1}{b^m}\sum_{\vx\in\mathcal{P}_m}f(\vx)}
\le \overbrace{\frac{\tS_{\ell,m}(f)\homega(m) \wcomega(m-\ell)}{1 - \homega(m-\ell) \wcomega(m-\ell)}}^{\epsilon_I:=}, \label{errbd}
\end{equation}
where one may increase $m$ until the error bound $\epsilon_I$ is small enough.

@@

We start by detailing our procedure in the form of an algorithm. Then, we discuss a possible improvement by considering a new estimator recently introduced in \cite{Owen} for the estimation of first-order indices.

@@

In both cases we test if the respective error bounds $\varepsilon_{\underline{S}_u}$ and $\varepsilon_{\overline{S}_u}$ are lower than the tolerance $\epsilon$. For Variant $A$ the stopping criterion writes: $ \forall u \in \mathcal{D}, \varepsilon_{\underline{S}_u} \leq \epsilon \text{ and } \varepsilon_{\overline{S}_u} \leq \epsilon$,
for Variant $B$ it writes:
$ \forall u \in \mathcal{D}, \varepsilon_{\underline{S}_u} \leq \epsilon.$

@@

In pratice, one does not know which are the small Sobol' indices. To overcome this issue, we propose the following alternative for Variant $A$. If at the end of the first iteration, the estimate $\widehat{\underline{S}}_u$, $u \in \mathcal{D}$, is smaller than a specified threshold, then at the next iteration estimator (\ref{first.order.est}) is switched for ``Correlation 2" and a third Sobol' sequence $\mathcal{P''}_{m}=\{\vz_i\}_{i=0}^{2^m-1}$ is constructed to obtain the corresponding evaluations $f({\vz}_{i,u}:{\vx}_{i,-u})$.
